---
title: "A predictive modeling case study"
weight: 1
tags: [parsnip, recipe, rsample, tune]
categories: [model fitting, tuning]
output:
  blogdown::html_page:
    toc: true
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
source(here::here("content/start/common.R"))
```

```{r load, include = FALSE, message = FALSE, warning = FALSE}
library(readr)
library(vip)
library(tidymodels)
pkgs <- c("tidymodels", "readr", "glmnet", "ranger", "vip")
theme_set(theme_bw() + theme(legend.position = "top"))
```

`r req_pkgs(pkgs)`

# Introduction

The previous _Getting Started_ articles have been focused on single tasks related to modeling. This example is a broader case study of building a predictive model. It uses all of the previous topics.  


# Data Spending

The data can be found on the `tidymodels.org` site: 

```{r hotel-import, message = FALSE}
library(tidymodels)
library(readr)

hotels <- 
  readr::read_csv('https://bit.ly/hotel_booking_data') %>%
  mutate_if(is.character, as.factor) 

dim(hotels)

table(hotels$children)/nrow(hotels)
```


```{r tr-te-split}
set.seed(35225)
splits <- initial_split(hotels, strata = children)

hotel_train <- training(splits)
hotel_test  <- testing(splits)
```


```{r validation-set}
set.seed(948)
val_set <- validation_split(hotel_train, strata = children, prop = 0.80)
val_set
```


# A first model: logistic regression

```{r logistic-features}
holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", 
              "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")

lr_recipe <- 
  recipe(children ~ ., data = hotel_train) %>% 
  step_other(country, threshold = 0.025) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = holidays) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
```

```{r logistic-model}
lr_mod <- 
  logistic_reg(penalty = tune(), mixture = tune()) %>% 
  set_engine("glmnet")

lr_workflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(lr_recipe)
```

```{r logistic-grid}
lr_reg_grid <- 
  expand.grid(mixture = c(0.10, 0.25, 0.5, 0.75, 1),
              penalty = 10^seq(-3, -1, length.out = 20))
```


```{r logistic-fit, cache = TRUE}
tune_ctrl <- control_grid(save_pred = TRUE)
roc_only <- metric_set(roc_auc)

lr_res <- 
  lr_workflow %>% 
  tune_grid(val_set,
            grid = lr_reg_grid,
            control = tune_ctrl,
            metrics = roc_only)
```

```{r logistic-results}
lr_res %>%
  collect_metrics() %>%
  mutate(mixture = format(mixture)) %>%
  ggplot(aes(x = penalty, y = mean, col = mixture)) +
  geom_point() +
  geom_line() +
  scale_x_log10() +
  scale_color_brewer(palette = "Set1")
```



```{r logistic-best}
lr_best <- 
  lr_res %>%
  collect_metrics() %>% 
  filter(mixture == 1) %>% 
  arrange(desc(mean)) %>% 
  select(penalty, mixture) %>% 
  slice(1)
lr_best
```


```{r logistic-roc-curve}
lr_auc <- 
  lr_res %>% 
  collect_predictions(parameters = lr_best) %>% 
  roc_curve(children, .pred_children) %>% 
  mutate(model = "Logistic Regression")

autoplot(lr_auc)
```


# Something more complex: random forest


```{r rf-features}
rf_recipe <- 
  recipe(children ~ ., data = hotel_train) %>% 
  step_date(arrival_date) %>% 
  step_holiday(arrival_date) %>% 
  step_rm(arrival_date) 
```

```{r num-cores}
cores <- parallel::detectCores()
cores
```


```{r rf-model}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("classification")

rf_workflow <- 
  workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(rf_recipe)
```

```{r rf-fit, cache = TRUE}
rf_res <- 
  rf_workflow %>% 
  tune_grid(val_set,
            grid = 25,
            control = tune_ctrl,
            metrics = roc_only)
```


```{r rf-results}
autoplot(rf_res)
```


```{r rf-best}
rf_best <- select_best(rf_res, metric = "roc_auc")
rf_best
```

```{r rf-roc-curve}
rf_auc <- 
  rf_res %>% 
  collect_predictions(parameters = rf_best) %>% 
  roc_curve(children, .pred_children) %>% 
  mutate(model = "Random Forest")

bind_rows(rf_auc, lr_auc) %>% 
  ggplot(aes(x = 1 - specificity, y = sensitivity, col = model)) + 
  geom_path() +
  geom_abline(lty = 2) + 
  coord_equal()
```

```{r rf-final, cache = TRUE}
rf_mod <- 
  rand_forest(mtry = 5, min_n = 2, trees = 1000) %>% 
  set_engine("ranger", num.threads = 20, importance = 'impurity') %>% 
  set_mode("classification")

rf_workflow <- 
  workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(rf_recipe)

rf_fit <- rf_workflow %>% last_fit(splits)
```


```{r rf-importance}
library(vip)

rf_fit$.workflow %>% 
  pluck(1) %>% 
  pull_workflow_fit() %>% 
  vip(num_features = 20) 
```

# Test set results

```{r test-set-roc-curve}
rf_fit %>% 
  collect_predictions() %>% 
  roc_auc(children, .pred_children)

rf_fit %>% 
  collect_predictions() %>% 
  roc_curve(children, .pred_children) %>% 
  autoplot()
```




# Session information

```{r si, echo = FALSE}
small_session(pkgs)
```


