---
title: List of parsnip models
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  digits = 3,
  collapse = TRUE,
  comment = "#>"
  )
options(digits = 3)
source("get_pkgdown_urls.R")
library(DT)
pkgs <- c("parsnip", "discrim")
library(discrim)
```

`parsnip` contains wrappers for a number of models. For example, the `parsnip` function `rand_forest()` can be used to create a random forest model. The **mode** of a model is related to its goal. Examples would be regression and classification. 

Since there are many models available in R, there are also "side packages" that wrap groups of models: 

 * [`discrim`](https://github.com/tidymodels/discrim) contains discriminant analysis models. 


```{r model-table, include = FALSE}
mod_names <- get_from_env("models")

model_info <-
  map_dfr(mod_names, ~ get_from_env(paste0(.x, "_predict")) %>% mutate(alias = .x))

model_modes <- 
  model_info %>% 
  distinct(alias, mode) %>% 
  group_by(alias) %>% 
  summarize(mode = paste0(sort(mode), collapse = ", "))

model_eng <- 
  model_info %>% 
  distinct(alias, engine) %>% 
  group_by(alias) %>% 
  summarize(engine = paste0("<tt>", sort(engine), "</tt>", collapse = ", "))

pkg_urls <- 
  get_pkgdown_urls(pkgs) %>% 
  dplyr::filter(alias %in% mod_names) %>% 
  mutate(title = str_remove(title, "General Interface for ")) %>% 
  full_join(model_modes, by = "alias") %>% 
  full_join(model_eng, by = "alias") %>% 
  dplyr::select(-alias)
```

_How_ the model is created is related to the **engine**. In many cases, this is an R modeling package. In others, it may be a connection to an external system (such as Spark or Tensorflow). The tables below list the engines for each model type. 

Models can be added by the user too. See the ["Making a {parsnip} model from scratch" vignette](https://tidymodels.github.io/parsnip/articles/articles/Scratch.html).


```{r table-display, echo = FALSE, results = "asis"}
DT::datatable(
  pkg_urls,
  rownames = FALSE,
  escape = FALSE,
  filter = "top",
  style = 'bootstrap',
  width = 800
)
```

Tidymodels use standardized parameter names to have some consistency between models and engines. The mapping between the `parsnip` arguments and their original names is: 

```{r params, echo = FALSE, results = "asis"}
envir <- parsnip::get_model_env()

args <- 
  ls(envir) %>% 
  tibble(name = .) %>% 
  filter(str_detect(name, "args")) %>% 
  mutate(model = str_replace(name, "_args", ""),
         args  = map(name, ~envir[[.x]])) %>% 
  unnest(args) %>% 
  select(model:original)

DT::datatable(args, filter = "top", style = 'bootstrap')
```


