---
title: List of Recipe Steps
---

```{r ex_setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  digits = 3,
  collapse = TRUE,
  comment = "#>"
  )
options(digits = 3)
script <- here::here("static/code/get_pkgdown_urls.R")
source(script)
library(DT)
pkgs <- c("recipes", "embed", "textrecipes", "themis")
```

There are many recipe steps in the `recipes` package. Other package that contain steps:

 * [`embed`](https://github.com/tidymodels/embed) contains methods for feature embeddings, effect encoding, UMAP and other methods. 

 * [`textrecipes`](https://github.com/tidymodels/textrecipes) has steps for natural language processing. 
 
 * [`themis`](https://github.com/tidymodels/themis) adds additional steps for dealing with class imbalances. 
 This table lists the functions in CRAN versions of the tidymodels packages. 

This table contains a list of steps and checking functions across these packages: 

```{r table-compute, include = FALSE}
pkg_urls <- get_pkgdown_urls(pkgs, fltr = "(^check_)|(^step_)")
```

```{r table-display, echo = FALSE, results = "asis"}
DT::datatable(
  pkg_urls,
  rownames = FALSE,
  escape = FALSE,
  filter = "top",
  style = 'bootstrap',
  width = 800
)
```
